\textsc{Does Copyright Affect Reuse?\\ Evidence from Wikipedia and the Google Books Digitization Project}

### 2016-09-13 writing/paper.tex
removed comments

% packages i use in many documents but leave off by default
\usepackage{amsmath, amsthm, amssymb}
%\usepackage{amsmath}
% \usepackage{dcolumn}
% \usepackage{endfloat}

@@

%\linespread{1.45}
\linespread{1.2}

@@

\title{Does Copyright Affect Reuse?\\ Evidence from the Google Books Digitization Project}
%\title{Does Copyright Limit Gains From Digitization?\\ Evidence from Google Books and Wikipedia}

%%\footnote{I am thankful to Erik Brynjolfsson, Daniel Fehder, Jeff Furman, Shane Greenstein, Joshua Krieger, Matt Marx, Aruna Ranganathan, Fabian Waldinger, Pai-Ling Yin, participants of the Economic Sociology Working Group and MIT Economics Third Year Lunch and especially Pierre Azoulay, Scott Stern, Catherine Tucker and Heidi Williams for comments. All errors remain mine.}

\author{Abhishek Nagaraj \footnote{MIT Sloan School of Management. E62-368, 100 Main Street, Cambridge, MA 02139. I am thankful to Erik Brynjolfsson, Daniel Fehder, Jeff Furman, Avi Goldfarb, Shane Greenstein, Joshua Krieger, Matt Marx, Aruna Ranganathan, Fabian Waldinger, Pai-Ling Yin, participants of the Economic Sociology Working Group and MIT Economics Third Year Lunch and especially Pierre Azoulay, Scott Stern, Catherine Tucker and Heidi Williams for comments. All errors remain mine.}
\\         \texttt{nagaraj@mit.edu}   
}

%%\date{\today}
\date{May 8, 2016}

@@

\begin{abstract}
\normalsize
While digitization projects like Google Books have dramatically increased access to information, this study examines how the ability to reuse digital information and diffuse knowledge to a wider audience critically depends on features of copyright law. I use the digitization of both copyrighted and non-copyrighted issues of \emph{Baseball Digest} by the Google Books Digitization Project to measure the impact of copyright on a prominent venue for reuse: Wikipedia. A specific feature of the 1909 Copyright Act, \emph{copyright renewal requirements} ensure that material published before 1964 is out of copyright, which allows for causal estimation of the impact of copyright on Wikipedia across this sharp cutoff. I find that digitization by Google Books increased the reuse of Baseball Digest on Wikipedia, but copyright hurt reuse by reducing citations to copyrighted issues by over 135\%. Copyright also had real impacts on readership -- affected pages had, on average, 20\% lower gains in internet traffic than pages unaffected by copyright. Finally, the impact of copyright on digitized material was highly uneven. Copyright mostly affected the reuse of images rather than text, and prevented reuse of information about less-popular players as compared to more popular ones.

%%I hypothesize and find that 
%%Copyright mostly prevented the reuse of images rather than text, and copyright significantly hurt the quality of Wikipedia pages for less-popular baseball players compared to their more popular counterparts.

%%However, the negative impact of copyright on diffusion is more pronounced for images compared to text and becomes large only when costs of access to digital material are reduced through the Google Books project. %%The main contribution of this paper is to provide the first causal estimates of the impact of copyright on reuse, a key measurement for copyright policy. 
\end{abstract}

%% I exploit a specific feature of the 1909 Copyright Act, \emph{copyright renewal} requirement, which caused issues of one magazine \emph{Baseball Digest} published before 1964 to fall in the public domain, while issues published after 1964 retained their copyright for this exercise. Specifically,

@@

The increasing digital representation of information has impacted a wide range of economic activities \citep{greenstein_digitization_2013}. Digitization has reduced the cost of storage, computation and transmission of information and enabled massive changes in the ways that creative producers build upon and reuse existing information  \citep{goldfarb_introduction_2014}. For example, the digitization of music has enabled new forms of production, remixing and reuse which has resulted in a new wave of innovation in the music industry \citep{waldfogel_digitization_2014}. Similarly, the digitization of ``prior art'' about traditional medicine has significantly affected new innovations in this area \citep{choudhury_ex-ante_2015}.

%Similarly, the availability of new digital maps of the earth's surface have led to a new wave of discoveries in the gold exploration industry \citep{nagaraj_private_2015}.

%There remains significant potential for the wider diffusion of digital information to transform markets in the near future.

@@

%%While diametrically opposed, these two theories are clear in their respective predictions; according to one theory, copyright lubricates the digital market for ideas, and, according to the other, it impedes the free diffusion of digital information.

%%Prospect theory \citep{kitch_nature_1977, hardin_tragedy_1968, landes_indefinitely_2002,liebowitz_seventeen_2004} holds that broad and strong intellectual property rights are useful, because they both incentivize the creation of new knowledge \citep{gans_remix_2014} and spur reuse by reducing contracting costs for the licensing of existing digital information by stimulating the maintenance of works already created (\cite{mazzoleni_economic_1998}, \cite{gallini_intellectual_2002}, \cite{gans_is_2010}, \cite{arora_markets_2001}). However, under a competing logic, copyright imposes transaction costs that prevent the exchange of information. These transaction costs are likely to be

%% Given contradictory theoretical predictions, systematic empirical analysis of the impact of copyright on reuse could help make progress on this question. Unfortunately, such analysis has largely been absent in the innovation and management literature on the topic, perhaps due to empirical challenges. Issues of data and measurement are central. Unlike scientific output which can be measured through patents and the publication record, the diffusion of cultural information is more informal and harder to track. When data can be located, copyright applies to all information by default and persists often for over a hundred years. This, makes it rather difficult to observe diffusion in the absence of copyright. Finally, even if these problems can be resolved, comparing reuse of copyrighted and non-copyrighted information is challenging because unobserved variables like the underlying quality of information are often correlated with reuse. Put another way, we do not know if information was reused because it was off-copyright or because it was inherently of higher quality.

%%Previous work in copyright has been largely theoretical and has focused on estimating the elasticity of new creative products with the strength of copyright protection (\cite{varian_copyright_2006}, \cite{watt_copyright_2006}). However, because digital  the creative process is cumulative in many markets (\cite{williams_intellectual_2013}, \cite{furman_climbing_2011}, \cite{scotchmer_standing_1991}), copyright may also affect reuse by changing the cost of reusing creative inputs in the production of new knowledge (\cite{li_dead_2012}, \cite{lerner_lost_2012}). This second impact is the focus of this study.

%% According to existing theory, whether copyright impedes or promotes the reuse of knowledge is unclear.

@@

%%is also relevant. This paper adds to this literature by focusing on the causal impact of copyright on \emph{reuse}, by using micro-data on knowledge flows and by separately identifying the interaction between digitization and copyright.

%% The findings that there are heterogeneous impacts on players of different qualities are interesting and relate to a few past studies: a. Napster helped concert revenue of less popular singers more than the established singers (Mortimer et. al. 2014), b. counterfeiting has a larger advertisement effect for the brands that were less well-known at the time of infringement and newer brands that had less sub-brands (Qian 2014).

%%In sum, I find that while digitization projects can help the broader diffusion of knowledge, copyright can have negative impacts on reuse. More broadly, this research shows how copyright policy can strongly influence the rate and direction of follow-on innovation on the internet. The present study joins the literature on digitization by pointing out the important role of intellectual property in influencing the impact of digital information on economic outcomes. Further, I also contribute more broadly, to the nascent empirical literature on copyright as well as related work on innovation, digitization and privacy. Of note is work in the legal literature on copyright that estimates the impact of copyright on availability and compares works produced before and after the US copyright cutoff date of 1923 \citep{heald_property_2007, heald_testing_2009, buccafusco_bad_2012}. Other work that studies the impact of copyright on prices and access in the market for historical books \citep{li_dead_2012, reimers_effects_2013} is also relevant. This paper adds to this literature by focusing on the causal impact of copyright on \emph{reuse}, by using micro-data on knowledge flows and by separately identifying the interaction between digitization and copyright. Finally, this paper contributes to a broader literature on knowledge flows (\cite{agrawal_gone_2006}, \cite{catalini_microgeography_2012}, \cite{singh_geographic_2013}) and the role of intellectual property (\cite{murray_formal_2007}, \cite{mowery_does_????}, \cite{hegde_tacit_2011}).

%%This study is also related to the literature on privacy \citep{goldfarb_privacy_2011} which focuses on the role of informational frictions in digital markets.

%%Having examined the overall impact of the copyright policy, I then turn to analyzing the differential impact of the program using the Wikipedia

%%: copyright significantly reduces the reuse of Baseball Digest material digitized by the Google Books project.

%Panel B, suggests that copyright law shaped this process in important ways.

%This figure tracks the total citations to Baseball Digest issu
%% The results show that the digitization of \emph{Baseball Digest} had a large positive impact on Wikipedia baseball pages, increasing the number of images by 140\% and text by about 50\% as compared to basketball pages. However, out-of-copyright pages benefit disproportionately compared to in-copyright pages, suggesting that copyright does indeed prevent reuse in this setting. Out-of-copyright pages receive a boost of about 80\% more images, and 140\% more citations in OLS specifications. However, such gains are unevenly distributed. The positive impact on reuse is many orders of magnitude larger for images than for text, suggesting that copyright is more effective in preventing the reuse of rich media like images and video than of text. Further, copyright seems to disproportionately harm players in the bottom 50 percentile of the quality distribution, suggesting that players who have few alternate sources of information benefit most from the release of out-of-copyright archival material. Finally, the difference in the amount of information on Wikipedia pages caused by the 1964 copyright experiment seems to have real impacts on Wikipedia's readership-- out-of-copyright pages receive a boost in monthly page-views of about 30-80\% as compared to in-copyright pages according to fixed-effects and instrumental variables specifications. A back-of-the-envelope calculation (see Appendix \ref{sec:envelope}) suggests losses to Wikipedia to the tune of about \$370,000 annually due to restrictions around the reuse of copyrighted material.

%%ask two additional questions about the data

%% I construct two different samples to understand the impact of copyright on the reuse of Baseball Digest content before and after its digitization. Sample A

%%all baseball players

%%Specifically, the estimation proceeds as follows. First, I identify a set of about five hundred prominent baseball players and a comparable set of basketball players (to act as an additional layer of control) who were active between 1944 and 1984. For this set of over 1000 players, I collect data from 13 different versions of their Wikipedia pages, one each from years 2001 to 2013. For each of these page-year observations, I count the number of images, the words of text, and citations to \emph{Baseball Digest}. I also collect data on internet traffic, which is available from 2007 to 2013. All players who made their debut before 1964 form the \emph{out-of-copyright} group because they could potentially benefit from uncopyrighted issues, while the rest are assigned to the \emph{in-copyright} group for analysis. The research design then estimates the impact of copyright on reuse by comparing the change in content for in-copyright and out-of-copyright baseball pages before and after the Google Books digitization event in a differences-in-differences (DD) framework. In other words, if out-of-copyright baseball pages see a larger increase in information after the digitization event than in-copyright pages after controlling for player and time-dependent fixed effects, we can attribute this difference to be the causal impact of copyright on the reuse of information. An additional level of analysis then uses DD estimates obtained separately from both baseball and basketball players to estimate a ``differences-in-differences-in-differences''(DDD) estimate of the impact of copyright on reuse that is robust to many alternate explanations.

%%The research design then proceeds in a differences-in-differences-in-difference (DDD) framework by constructing three levels of differences that provides me with an estimate of the impact of copyright on reuse in Wikipedia that is robust to a number of different alternative explanations. Specifically, I first construct differences-in-differences estimates that compare the change in content between In-Copyright pages (player debuts after 1964) and Out-of-Copyright pages (player debut before 1964) for both baseball players and basketball players separately, and then compare these two estimates to arrive at a DDD estimate of the impact of copyright on reuse.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Digitization and the Role of Copyright}
%\label{sec:theory}
%\subsection{Digitization and Reuse}
%\subsection{How Copyright could Affect Reuse}

@@

%%\subsubsection*{Google Books and the Baseball Digest Copyright Experiment}

@@

%of over 200,000 periodicals in WorldCat (a knowledge repository), only about 1300 serials had renewed copyrights for a second 28-year copyright term before 1964.
Although copyrights for some movies and books were renewed successfully, because this requirement was not well-known, it \emph{``tripped up many smaller publishers, and a failure to renew caused many works to lapse into the public domain''} \citep{andrade_copyright_2014}.  According to research by the University of Pennsylvania library system \citep{mark_ockerbloom_next_2006}, a failure to renew copyrights has meant that issues of many periodicals published before 1964, have fallen into the public domain including issues of journals like the \emph{American Economic Review}, \emph{Bell System Technical Journal}, \emph{Biometrika} and periodicals like \emph{The Baltimore Sun}, \emph{The Los Angeles Evening Herald}, \emph{The Kansas City Star} etc.\footnote{While a number of different publications did not renew copyright, I focus on Baseball Digest, because it was one of the few that was also digitized in its entirety by the Google Books project.}

@@

 %%Baseball players in the sample have played an average of 1280 games in their career, made their debut around 1966 and played for an average of 14.70 years in their career.

% In addition to data on prominent baseball players, I also collected similar data on a comparable group of basketball players. While the core of the analysis focuses on analyzing the impact of copyright within the set of baseball players, adding another layer of controls using basketball player data helps strengthen my analysis. This data allows estimation of the baseline impact of digitization, in addition to providing additional controls that can account for overall dynamics on the Wikipedia platform. Basketball is a convenient choice because like baseball, basketball also has an active community of editors on Wikipedia. However, basketball did not experience any digitization event like the availability of \emph{Baseball Digest} through Google Books. For basketball, I obtained data from \url{databasebasketball.com} which provides names, years of debut and other performance data for individual players. To identify a set of players comparable to the baseball players, I used the dataset for the Top 1000 players by career minutes played and choose those who made their debut between 1944 and 1984.\footnote{Basketball Hall of Fame nominations data is not feasible because the voting process does not begin until 1959} 
%%The 564 Basketball players in the sample played about 716 games, scored 9223 points over their career and made their debut around 1970.

@@

%% In other words, Sample A tracks citations to publication-year 1963 by track how many times \emph{Baseball Digest (1963)} was cited on 2004 versions of Wikipedia pages, on 2005 versions of Wikipedia pages and so on, till 2012.

%%Using these data, I build Sample A, a panel of citation counts at the publication-year level for 2004-2012.

@@

%, Images, Text} and \emph{Traffic}. The data show that baseball Wikipedia pages contain an average of 0.15 citations, 0.56 images and 969 words and that they receive on average 105 page views per month. For basketball players, Wikipedia pages have an average of 0 citations to \emph{Baseball Digest}, 0.19 images and 540 words and receive an average of 123 page views per month.
\subsubsection{Sample B: Player-Page Level}

@@

To build this sample, I first used the ``Baseball Hall of Fame'' voting dataset by Sean Lahman\footnote{see \url{http://www.seanlahman.com/baseball-archive/statistics/}} to compile a list of 541 players who have been nominated for election to the Baseball Hall of Fame and who made their debut appearances between 1944 and 1984. The Hall of Fame nomination list allowed me to include players who had finished their careers and who had passed a screening committee judgment, but it also ``removes from consideration players of clearly less qualification'' \citep{abbott_future_2011}. Thus, the nomination list can be said to include only those players who merit encyclopedic inclusion. The dataset also provides biographical details of the players including date of debut and performance details like their experience, length of career and number of appearances in all-star games.  
%and the length of their playing career.

%%\footnote{December 1 is a convenient date, because it provides measurements in yearly intervals from the digitization event which happened on December 9. 2001 was the year when Wikipedia began and 2012 is the final year of analysis.}

@@

%%For basketball players the \emph{quality} rank is assigned based on the number of career minutes played.

@@

%% how did digitization help?
%% what was the role of copyright?
%% -- cross section + killer picture
%% -- DD results
%% check trends, robustness
%% differential -- theory
%% differential: images, text
%% differential: player quality (B)
%% traffic (B)

%% to discuss: summary
%% Tables
%% summary stats (A/B)
%% reuse outcomes (A/B)
%% DD (sample A)
%% DD (sample B)
%% traffic (B)
%%
%% Figures
%% schematic
%% digitization / separate
%% killer (sample A)
%% DD time (A)
%% killer + DD (B)
%% traffic scatter
%% hetero

@@

%% 3.1 simple descriptives
%% Fig 2
%% Table 2

@@

%In this section, I investigate the impact of the Google Books copyright experiment on affecting the reuse of magazine material on Wikipedia.

@@

%Taken together this analysis suggests that within a subset of Wikipedia pages belonging to prominent baseball players, reuse of the Baseball Digest magazine was significantly lower for copyrighted material. 
Taken together, estimates from both Sample A and Sample B are able to robustly confirm the main hypothesis of this study--the copyright restrictions on digitized, post-1964 Baseball Digest material significantly reduced the likelihood of reuse on Wikipedia.

@@

%The empirical evidence presented so far, provides support for the positive effects on reuse from digitization, and the negative effects of copyright on the reuse of digital material.

@@

%When no IP is asserted and material is in the public domain, the reduction in access costs is the greatest. However, when material is digitally available but IP rights need to be negotiated, the reduction in access costs due to digitization is muted. This is because, IP introduces transaction costs which make it more difficult to reuse digital material. This characterization of IP is consistent with existing work in the innovation literature that finds that when scientific material is patented, transaction costs have a negative effect on follow-on innovation.

%Therefore, when transaction costs due to IP are considered, the positive effect of digitization due to reduced access costs depends on the copyright status of digitized material. Specifically, the benefits of digitization on reuse are the highest for material that are not protected by copyright because of zero transaction costs, while for in-copyright material, reuse after digitization is substantially lower. These predictions are consistent with Fig \ref{fig:overview} Panel (2).

%\subsubsection{The Role of Access Costs and Transaction Costs}

@@

%%Having clarified the basic channel through which digitization and copyright affect the overall level of reuse, I now hypothesize how these two factors could have distributional consequences as well. I focus on two different margins: first, the differential impact of IP on the reuse of images vs. text, and for Wikipedia topics of different inherent quality.

@@

%reuse, even with significant modifications could constitute infringement .
Consequently, in order to prevent infringement, the reuse of textual information with sufficient modifications is typically possible at low costs and relatively simple. However, for images to be legally reused, large modifications need to be made to satisfy the ``transformative reuse'' criteria, making the process more complicated and significantly costly. In practice, even when significant changes are made, copyright infringement is possible (see \emph{Cariou v. Prince, 714 F.3d 694 (2d Cir. 2013)} for an example) and, therefore, end-users often avoid the reuse of copyrighted images without explicit permission from the creators.

@@

%In light of the theoretical framework, the effects of different levels of transaction costs for images vs. text, will lead to distributional effects of IP by the type of underlying media. Digital images are significantly more likely to be affected by copyright as compared to digitized text. This section will test this hypothesis in greater detail.

% For text, transformation usually occurs through paraphrasing, i.e. rewording the basic facts in different language. Given that factual material is exempt from copyright, such paraphrasing is quite common in practice even when the underlying material is copyright. On the other hand, for images to be reused, there must be significant changes for the image to be reused and not constitute copyright infringement.

@@

%% papers to cite: (Mortimer et. al. 2014), Qian 2014
%% brynolsffson: long tail, goldfarb papers, bitnet paper, catalini -- crowdfunding, waldfogel paper, 
%% kevin bryan: open access literature
%% IP literature: williams, stern and murray on IP

@@

%%This is very similar to a ``democratization'' effect or ``long tail'' effect hypothesized in the literature.
In order to understand how copyright affects player-pages, I argue that the optimal level of knowledge on Wikipedia pages depends both on the value of new information to a page and the cost of adding new information. A large literature in media economics finds that the provision of news about events is directly proportional to commercial interest \citep{prat_political_2011, stromberg_natural_2007}. In line with this research, we can expect value of information about players to be directly proportional to player quality, while the costs of sourcing information to be inversely related. Higher quality players attract higher interest from end users and therefore the value of information for these players is greater. On the contrary, there are a number of alternate sources covering higher quality players which makes it cheaper to source information about them. The assumption for lower quality players is exactly the opposite: it is more expensive to source information and the value of such information is also lower. In this framework, given the costs and benefits of adding new information, there are greater incentives to add information for players of higher quality and this information is easier to obtain even before the digitization effort.

%%(which is increasing in player quality) and the cost of adding new information (which is inversely proportional to player quality).
Now, consider the reduction in the cost of access to information due to digitization \citep{goldfarb_introduction_2014} and increased transaction costs due to intellectual property \citep{williams_intellectual_2013}. For players of the highest quality, information already exists even before digitization, and the marginal utility of new information is low. For obscure players, even if information exists on \emph{Baseball Digest}, reuse is unlikely given the low value of adding information and the fact that these players are rarely featured on Wikipedia to begin with. In this framework, the value of out-of-copyright information is highest for a middle tier of player quality: players who are good enough to merit encyclopedic inclusion, but information about whom was relatively difficult to source before digitization. Therefore the prediction is that, out-of-copyright digital information is most valuable for player-pages at the middle tier of the quality distribution, rather than for players at the very top. Given that the lowest tier of player quality is unlikely to be covered on Wikipedia, within the sample of about 500 players who have been nominated for the Hall of Fame, the bottom two quartiles are likely to represent this ``middle tier.''

@@

%% papers to cite: (Mortimer et. al. 2014), Qian 2014
%% brynolsffson: long tail, goldfarb papers, bitnet paper, catalini -- crowdfunding, waldfogel paper, 
%% kevin bryan: open access literature
%% IP literature: williams, stern and murray on IP

%%In this setup, players for which the digitized information is most useful are players who are ``good enough'' (i.e. the middle tier of player quality). Players of higher quality already have higher levels of information before the digitization event and, therefore, the out-of-copyright information is less valuable. For players at the lowest level of quality, the value of adding out-of-copyright information is lower and, therefore, the digitization event is less valuable. Appendix \ref{sec:model} presents a formal ``toy'' model that formalizes this idea and delivers the basic insight: out-of-copyright digital information is most valuable for player-pages at the middle tier of the quality distribution, rather than for players at the very top.

%% 
%% copyright hurts reuse
%% why? lowers the cost of reuse
%% 
\subsection{Comparing the Reuse of Images vs. Text}
\label{sec:diff1}

@@

%% Further, following Table \ref{tab:maindd}, I leverage Sample B to evaluate the impact of the copyright cutoff on the level of information on Wikipedia pages, by looking separately at image and text content. This analysis helps us understand the role of copyright in differentially affecting different margins of information for a single Wikipedia page.

@@

%%the evidence in Figure \ref{fig:quality} is consistent with the theory that out-of-copyright material is most beneficial to players who are notable enough to be included in the encyclopedia, but not famous enough to have substitute information about them available in other sources.

@@

%% IP literature

The results are also related to another stream of work on the empirical effects of intellectual property on the diffusion of knowledge. Some studies \citep{murray_formal_2007, murray_mice_2009, williams_intellectual_2013, furman_climbing_2011} find a generally negative effect of intellectual property on follow-on use, while more recent evidence seems to be mixed (\cite{sampat_how_2014}, \cite{galasso_patents_2015}). This study provides direct evidence to this literature that the costs of access (i.e. digitization) seem to matter for the impacts of IP on reuse. From a policy point of view, this paper is able to directly address questions that are likely to be important going forward such as: (a) how does the impact of copyright change when works are digitized and access costs are low, and (b) does copyright need to be modified for the digital age?

%% managerial implications

Finally, this study also has implications for managers in knowledge-intensive sectors of the economy. For those in-charge of IP and digitization strategy, this study suggests that copyright can be an effective intellectual property tool to manage digital assets. How effective copyright can be seems to depend on access and the medium in which information is expressed. This is useful because there is often a concern that piracy is so rampant on the internet that tools other than traditional intellectual property (like DRM) are often necessary \citep{zhang_intellectual_2014} to supplement toothless copyright law. Second, for managers who are interested in using user communities like Wikipedia to generate innovation (\cite{boudreau_incentives_2011}, \cite{franzoni_crowd_2014}) or open innovation more broadly \citep{fleming_brokerage_2007}, this study suggests that provision of external, uncopyrighted but digitized material can be extremely beneficial. This study finds that knowledge within user communities is often sourced from external sources, and policy measures that affect the availability and legal status of sources can either boost or retard innovative activity within such communities.

@@

%%(specifically the impact of copyright on the incentives to create new content is not examined), it is still useful, because it helps estimate what potential incentives copyright needs to provide to creators in order to justify the losses estimated to society.

%\newpage
\baselineskip 14.2pt
\bibliography{baseball}
\bibliographystyle{chicago}

%\section{Tables and Figures}

%%% INSERT TABLES
%\input{tables.tex}
\input{tables.tex}
%\input{/home/nagaraj/Dropbox/research/copyright/writing/table_include.tex}

>>>>

% packages i use in many documents but leave off by default
\usepackage{amsmath, amsthm, amssymb}

@@

\linespread{1.2}

@@

\title{\textsc{Does Copyright Affect Reuse?\\ Evidence from Wikipedia and the Google Books Digitization Project}}

\author{Abhishek Nagaraj \footnote{Haas School of Business, University of California, Berkeley. F565, 2220 Piedmont Ave, Berkeley, CA 94720-1900. I am thankful to Erik Brynjolfsson, Daniel Fehder, Jeff Furman, Avi Goldfarb, Shane Greenstein, Joshua Krieger, Matt Marx, Aruna Ranganathan, Fabian Waldinger, Pai-Ling Yin, participants of the Economic Sociology Working Group and MIT Economics Third Year Lunch and especially Pierre Azoulay, Scott Stern, Catherine Tucker and Heidi Williams for comments. All errors remain mine.}
\\         \texttt{nagaraj@berkeley.edu}   
}

\date{\today}

@@

\begin{abstract}
\normalsize
While digitization projects like Google Books have dramatically increased access to information, this study examines how the ability to reuse digital information and diffuse knowledge to a wider audience critically depends on features of copyright law. I use the digitization of both copyrighted and non-copyrighted issues of \emph{Baseball Digest} by the Google Books Digitization Project to measure the impact of copyright on a prominent venue for reuse: Wikipedia. A specific feature of the 1909 Copyright Act, \emph{copyright renewal requirements} ensure that material published before 1964 is out of copyright, which allows for causal estimation of the impact of copyright on Wikipedia across this sharp cutoff. I find that digitization by Google Books increased the reuse of Baseball Digest on Wikipedia, but copyright hurt reuse by reducing citations to copyrighted issues by over 135\%. Copyright also had real impacts on readership -- affected pages had, on average, 20\% lower gains in internet traffic than pages unaffected by copyright. Finally, the impact of copyright on digitized material was highly uneven. Copyright mostly affected the reuse of images rather than text, and prevented reuse of information about less-popular players as compared to more popular ones. 
\end{abstract}

@@

The increasing digital representation of information has impacted a wide range of economic activities \citep{greenstein_digitization_2013}. Digitization has reduced the cost of storage, computation and transmission of information and enabled massive changes in the ways that creative producers build upon and reuse existing information  \citep{goldfarb_introduction_2014}. For example, the digitization of music has enabled new forms of production, remixing and reuse which has resulted in a new wave of innovation in the music industry \citep{waldfogel_digitization_2014}. Similarly, the digitization of ``prior art'' about traditional medicine has significantly affected new innovations in this area \citep{choudhury_ex-ante_2015}. Similarly, the availability of new digital maps of the earth's surface have led to a new wave of discoveries in the gold exploration industry \citep{nagaraj_private_2015}.

@@

@@

@@

@@

Although copyrights for some movies and books were renewed successfully, because this requirement was not well-known, it \emph{``tripped up many smaller publishers, and a failure to renew caused many works to lapse into the public domain''} \citep{andrade_copyright_2014}.  According to research by the University of Pennsylvania library system \citep{mark_ockerbloom_next_2006}, a failure to renew copyrights has meant that issues of many periodicals published before 1964, have fallen into the public domain including issues of journals like the \emph{American Economic Review}, \emph{Bell System Technical Journal}, \emph{Biometrika} and periodicals like \emph{The Baltimore Sun}, \emph{The Los Angeles Evening Herald}, \emph{The Kansas City Star} etc.\footnote{While a number of different publications did not renew copyright, I focus on Baseball Digest, because it was one of the few that was also digitized in its entirety by the Google Books project.}

@@

@@

@@

\subsubsection{Sample B: Player-Page Level}

@@

To build this sample, I first used the ``Baseball Hall of Fame'' voting dataset by Sean Lahman\footnote{see \url{http://www.seanlahman.com/baseball-archive/statistics/}} to compile a list of 541 players who have been nominated for election to the Baseball Hall of Fame and who made their debut appearances between 1944 and 1984. The Hall of Fame nomination list allowed me to include players who had finished their careers and who had passed a screening committee judgment, but it also ``removes from consideration players of clearly less qualification'' \citep{abbott_future_2011}. Thus, the nomination list can be said to include only those players who merit encyclopedic inclusion. The dataset also provides biographical details of the players including date of debut and performance details like their experience, length of career and number of appearances in all-star games.

@@

@@

@@

%% structure
%% 3.1 simple descriptives
%% Fig 2
%% Table 2

@@

@@

Taken together, estimates from both Sample A and Sample B are able to robustly confirm the main hypothesis of this study--the copyright restrictions on digitized, post-1964 Baseball Digest material significantly reduced the likelihood of reuse on Wikipedia.

@@

@@

@@

@@

Consequently, in order to prevent infringement, the reuse of textual information with sufficient modifications is typically possible at low costs and relatively simple. However, for images to be legally reused, large modifications need to be made to satisfy the ``transformative reuse'' criteria, making the process more complicated and significantly costly. In practice, even when significant changes are made, copyright infringement is possible (see \emph{Cariou v. Prince, 714 F.3d 694 (2d Cir. 2013)} for an example) and, therefore, end-users often avoid the reuse of copyrighted images without explicit permission from the creators.

@@

@@

@@

In order to understand how copyright affects player-pages, I argue that the optimal level of knowledge on Wikipedia pages depends both on the value of new information to a page and the cost of adding new information. A large literature in media economics finds that the provision of news about events is directly proportional to commercial interest \citep{prat_political_2011, stromberg_natural_2007}. In line with this research, we can expect value of information about players to be directly proportional to player quality, while the costs of sourcing information to be inversely related. Higher quality players attract higher interest from end users and therefore the value of information for these players is greater. On the contrary, there are a number of alternate sources covering higher quality players which makes it cheaper to source information about them. The assumption for lower quality players is exactly the opposite: it is more expensive to source information and the value of such information is also lower. In this framework, given the costs and benefits of adding new information, there are greater incentives to add information for players of higher quality and this information is easier to obtain even before the digitization effort.

Now, consider the reduction in the cost of access to information due to digitization \citep{goldfarb_introduction_2014} and increased transaction costs due to intellectual property \citep{williams_intellectual_2013}. For players of the highest quality, information already exists even before digitization, and the marginal utility of new information is low. For obscure players, even if information exists on \emph{Baseball Digest}, reuse is unlikely given the low value of adding information and the fact that these players are rarely featured on Wikipedia to begin with. In this framework, the value of out-of-copyright information is highest for a middle tier of player quality: players who are good enough to merit encyclopedic inclusion, but information about whom was relatively difficult to source before digitization. Therefore the prediction is that, out-of-copyright digital information is most valuable for player-pages at the middle tier of the quality distribution, rather than for players at the very top. Given that the lowest tier of player quality is unlikely to be covered on Wikipedia, within the sample of about 500 players who have been nominated for the Hall of Fame, the bottom two quartiles are likely to represent this ``middle tier.''

@@

\subsection{Comparing the Reuse of Images vs. Text}
\label{sec:diff1}

@@

@@

@@

%% IP literature
The results are also related to another stream of work on the empirical effects of intellectual property on the diffusion of knowledge. Some studies \citep{murray_formal_2007, murray_mice_2009, williams_intellectual_2013, furman_climbing_2011} find a generally negative effect of intellectual property on follow-on use, while more recent evidence seems to be mixed (\cite{sampat_how_2014}, \cite{galasso_patents_2015}). This study provides direct evidence to this literature that the costs of access (i.e. digitization) seem to matter for the impacts of IP on reuse. From a policy point of view, this paper is able to directly address questions that are likely to be important going forward such as: (a) how does the impact of copyright change when works are digitized and access costs are low, and (b) does copyright need to be modified for the digital age?

%% managerial implications
Finally, this study also has implications for managers in knowledge-intensive sectors of the economy. For those in-charge of IP and digitization strategy, this study suggests that copyright can be an effective intellectual property tool to manage digital assets. How effective copyright can be seems to depend on access and the medium in which information is expressed. This is useful because there is often a concern that piracy is so rampant on the internet that tools other than traditional intellectual property (like DRM) are often necessary \citep{zhang_intellectual_2014} to supplement toothless copyright law. Second, for managers who are interested in using user communities like Wikipedia to generate innovation (\cite{boudreau_incentives_2011}, \cite{franzoni_crowd_2014}) or open innovation more broadly \citep{fleming_brokerage_2007}, this study suggests that provision of external, uncopyrighted but digitized material can be extremely beneficial. This study finds that knowledge within user communities is often sourced from external sources, and policy measures that affect the availability and legal status of sources can either boost or retard innovative activity within such communities.

@@

\baselineskip 14.2pt
\bibliography{baseball}
\bibliographystyle{chicago}

%%% INSERT TABLES
\input{tables.tex}
